On the solution of katabatic flows with spatially varying eddy viscosity and diffusivity

### 2016-06-23 paper_BLM.tex
pre-review

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2014/09/25
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{natbib}
\linenumbers
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{epstopdf}
%\graphicspath{ {./images/results} }
\usepackage{color}
\usepackage{ctable}
%\usepackage{epsfig}
\usepackage{hyperref}
\usepackage{soul}
\usepackage{amssymb}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
 \journalname{Boundary-Layer Meteorology}

@@

\title{Exact analytic solution for slope flows with spatially varying eddy viscosity and diffusivity}

@@

%\authorrunning{Short form of author list} % if too long for running head
\author{M.~G.~Giometto \and R.~Grandi\and J.~Fang \and M.B. Parlange}

\institute{ M.G. Giometto \at School of Architecture, Civil and Environmental Engineering, \'Ecole Polytechnique F\'ed\'erale de Lausanne, Lausanne, Switzerland \\
		\email{marco.giometto@epfl.ch}
		\and
		Riccardo Grandi \at Mathematics Institute for Geometry and Applications, \'Ecole Polytechnique F\'ed\'erale de Lausanne, Lausanne, Switzerland \\
		 \and 
		 J. Fang \at School of Architecture, Civil and Environmental Engineering, \'Ecole Polytechnique F\'ed\'erale de Lausanne, Lausanne, Switzerland \\
		 \and
		M.B. Parlange \at Civil Engineering, Faculty of Applied Sciences, University of British Columbia, Vancouver, BC, Canada
}

@@

% ABSTRACT
\begin{abstract}
An exact analytic solution of the steady-state Prandtl model equations is derived, valid for spatially varying eddy diffusivities (O'Briens type) and Prandtl number of unity. 
In this formulation profiles show significant variations in both phase and amplitude of minima-maxima with respect to the classic constant eddy viscosity model and the more recent, approximate, WKB solution. 
The near wall region is characterized by a relatively stronger surface inversion and velocity gradients, the low-level-jet is further displaced toward the wall, and its peak velocity strongly depends on the model parameter, suggesting a tighter coupling between dynamics and thermodynamics.
\end{abstract}

\section{Introduction}
Natural convection of turbulent stratified flows over sloping surfaces is ubiquitous in nature and engineering, and is of interest not only as a fundamental problem in itself, but also because of the important role it plays over a broad range of scales and applications. 
On a local scale, for instance, it governs local weather conditions, affecting atmospheric transport processes and dispersion of scalars such as heat and humidity \citep{Monti2002, Nylen2004b}. 
On a regional scale, it drives the development of deep-convective circulations \citep{Egger1985, Parish1998} and is responsible for intense cyclonic vorticity in the middle and upper troposphere \citep{Parish1992}.
Further, recent experimental field campaigns \citep{Chu1987, Oerlemans1993} have shown that persistent katabatic winds characterize the atmospheric boundary layer over snow-ice surfaces and glaciers, and therefore an accurate characterization of such flows is an essential component of understanding and modeling of the weather and climate.
However, the complex dynamics (e.g. occurrence of intermittency, waves, Kelvin-Helmholtz instabilities and low-level jets) and the lack of a satisfactory similarity theory for such flows \citep{Nadeau2012} pose a heavy burden in terms of computational requirements for numerical modellers; in most cases the required resolution is prohibitively costly \citep{Shapiro2004, Shapiro2004a, Fedorovich2009, Fedorovich2009d, Burkholder2011a}.
Because of this, conceptual models are still of great interest, and represent a valid tool for the characterization of such systems. 
A cornerstone in the understanding of natural convection of stratified fluid along a sloping surface is represented by the classic Prandtl analytic model \citep{Prandtl}, and its more recent extensions/generalizations to include the effects of Coriolis force \citep{Gutman1964}, external winds \citep{Lykosov_1972}, surface heterogeneity \citep{Egger1985, Shapiro2007, Oldroyd2014}.
The Prandtl model approximates the atmosphere in a Boussinesq sense and describes a steady flow over a thermally perturbed unbounded planar sloping surface that lies within a stratified environment, given by the following system of equations:
%
\begin{linenomath*}
\begin{eqnarray}
	N^2 u(z) \sin{\alpha} & = & d( K_H db/dz )/dz,
	\label{eqn1u} \\
	%
	b(z) \sin{\alpha} & = & -d( K_M du/dz )/dz,
	\label{eqn1b}
\end{eqnarray}
\end{linenomath*}
%
where $z$ denotes the normal-to-slope coordinate directions, $u$ is the downslope velocity, $b$ is buoyancy, $N$ is the buoyancy frequency characterizing the system (related to the background stratification), $\alpha$ is the slope angle and $K_M$ and $K_H$ denote the eddy viscosity and diffusivity (an eddy viscosity/diffusivity model has been used to parametrize turbulent fluxes of momentum and buoyancy).
Equations are integrated in $z \in \ [0,H]$ with boundary conditions $u(0)=0$, $u(H) = 0$, $b(0)=b_s$ and $b(H) = 0$ ($b_s > 0$ for upslope flows, whereas $b_s<0$ for downslope flows).

The flow is assumed to be invariant in the downslope direction and the model can be used to determine its vertical structure \citep{NAPPO1987}.
Eqs. \ref{eqn1u} and \ref{eqn1b} state that downslope (upslope) convection of cold (warm) air is balanced by momentum flux divergence, and that along-slope generated buoyancy is balanced by buoyancy flux divergence; thus the model is applicable away from ridges and valleys, so that advection terms are negligible \citep{NAPPO1987}.
The Prandtl constant-K model is known to be overdissipative in the near surface regions, and therefore not able to represent the observed strong surface gradients of temperature and momentum that commonly characterize katabatic and anabatic flows \citep{Oerlemans1998, grisogono2001katabatic}. 
Simple variations in the eddy diffusivity profiles were introduced in an exact analytic solution by \citet{gutman_1983}, and more recently \citet{grisogono2001katabatic} considered general variations in the vertical structure of the eddy diffusivities, and derived  solutions based on the WKB approximation.
The WKB solution is able to account for additional dynamics while still retaining an elegant form.
However, that theory is only applicable when the model parameters $(K_M, K_H)$ vary more slowly than the solution $(u,b)$, and the validity of such a constraint for slope flows has been the subject of great debate \citep{Grisogono2002}.
In the following, we derive a closed-form solution to the Prandtl-model equations, valid for O'Brien's type \citep{O'Brien1970} eddy diffusivities and representing an exact alternative to the WKB solution -- for the chosen form of the model parameter.

% ###################################################################
\section{The Analytic Solution}
%
We consider the one-dimensional Prandtl model equations \ref{eqn1u} and \ref{eqn1b}. Assuming $Pr=1$ (i.e. $K_M=K_H=K$) and assigning length, velocity and buoyancy scales as $L=\max{(K)}^{1/2} N^{-1/2} \sin^{-1/2}{\alpha}$, $U=b_s N^{-1}$ and $B=b_s$ Eqs. \ref{eqn1u} and \ref{eqn1b} can be reduced to:
%
\begin{linenomath*}
\begin{subequations}
\label{ode}
\begin{align}
	u^+& = d( K^+ db^+/dz^+ )/dz^+, && \mathrm{in}  \ [0,H^+]
	\label{eqn1u_norm} \\
	%
	b^+ & = -d( K^+ du^+/dz^+ )/dz^+, &&\mathrm{in}  \ [0,H^+]
	\label{eqn1b_norm}
\end{align}
\end{subequations}
\end{linenomath*}
%
respectively, where $K^+$ is a normalized eddy viscosity, $z^+=zL^{-1}$, $u^+ = uU^{-1}$, $b^+=b B^{-1}$, with boundary conditions $u^+(0)=0$, $u^+(H^+) = 0$, $b^+(0)=\pm 1$ and $b^+(H^+) = 0$.
%
Further, we introduce the following variables 
%
\begin{linenomath*}
\begin{equation}
   f_1= b^+ - i u^+ \hspace{3mm} \textrm{and} \hspace{3mm} f_2=b^+ + iu^+,
   \label{combinationVars}
\end{equation}
\end{linenomath*}
%
and upon substitution of Eqs. \ref{eqn1u_norm} and \ref{eqn1b_norm} into Eqs. \ref{combinationVars}, the system can be decoupled to obtain two (independent) complex ordinary differential equations (ODEs) for the canonical variables $f_1$ and $f_2$:
%
\begin{linenomath*}
\begin{subequations}
\label{ode_norm}
\begin{align}
	i \cdot f_1 &=  d(  K^+ df_1/dz^+ )/dz^+, && \mathrm{in}  \ [0,H^+]
	\label{ode1} \\
	i \cdot f_2 &= - d( K^+ df_2/dz^+ )/dz^+, && \mathrm{in}  \ [0,H^+]
	\label{ode2}
\end{align}
\end{subequations}
\end{linenomath*}
%
with boundary conditions $f_i(0)=\pm 1$ and $f_i(H^+) = 0$, $i=1,2$.

We assume eddy diffusivities in line with the classic O'Brien's type \citep{O'Brien1970}, viz. $K^+ = A^+ (z^++\epsilon^+)(z^+-H^+-\epsilon^+)^2$, where
$\epsilon^+$ is an additional parameter that we introduce in order to shift the zeroes of $K^+$ outside the interval of integration $[0,H^+]$.
$\epsilon^+$ can be tuned to provide the desired value of viscosity / diffusivity at the wall.

In order to solve Eqs. \ref{ode1} and \ref{ode2}, we observe that both have the form
%
\begin{linenomath*}
\begin{equation*}
	f= \mp i \cdot d/dz^+ ( K^+ df/dz^+ ),
\end{equation*}
\end{linenomath*}
%
and therefore can be solved using the same method. First, we rewrite the equation in canonical form:
%
\begin{linenomath*}
\begin{align}
	\label{can_form}
	f''+P f'+ Qf &=0, && \mathrm{in}  \ [0,H^+]
\end{align}
\end{linenomath*}
%
where $P(z^+)=(K^+)'/K^+$ and $Q(z^+)=1/(\pm i K^+)$. Second, it is an easy computation to show that, if $f$ solves Eq. \ref{can_form} then $g(z^+):=f( (H^++2 \epsilon^+) z^+- \epsilon^+)$ solves:
%
\begin{linenomath*}
\begin{align}
	\label{can_form_sec}
	g''+\widetilde{P} g'+ \widetilde{Q} g =0, && \mathrm{in}  \ [0,1]
\end{align}
\end{linenomath*}
%
where $\widetilde{P}(z^+)=\gamma'(z^+)/\gamma(z^+)$ and $\widetilde{Q}(z^+)=1/(\pm i \cdot A^+ \cdot (H^++2\epsilon^+) \gamma(z^+))$, with $\gamma(z^+)=z^+(z^+-1)^2$.
Eqn. \ref{can_form_sec} is a second order ODE with three regular singular points at $z^+ = 0,1$ and $\infty$, as in \citet{Morse1954a}. 
This special case is known as the \emph{equation of Papperitz} and its general solution is:
%
%\begin{widetext}
\begin{linenomath*}
\begin{equation}
	g(z^+) = \alpha (1-z^+)^\mu \ _2F_1 (\mu, 1- \mu', 1+\mu-\mu', 1-z^+)
	+\beta (1-z^+)^{\mu'} \ _2F_1 (\mu', 1- \mu, 1+\mu'-\mu, 1-z^+),
	\label{gen_sol_g}
\end{equation}
\end{linenomath*}
%\end{widetext}
%
where $\ _2F_1$ are \emph{Gaussian hypergeometric functions} and $\mu$, $\mu'$ are the solutions to the degree two equation
%
\begin{linenomath*}
\begin{equation*}
	x^2+x+1/(\pm i A^+(H^++2 \epsilon^+))=0,
\end{equation*}
\end{linenomath*}
%
depending on which of Eqs. \ref{can_form_sec} we are solving.
Note that the general solution to Eq. \ref{can_form_sec} is also often written using the \emph{Riemann symbol}:
%
\begin{linenomath*}
\begin{equation*}
\left \{
\begin{array}{cccc}
0 & 1 & \infty &  \\
\lambda & \mu & \nu & z^+\\
\lambda' & \mu' & \nu' & \\
\end{array}
\right \}
\end{equation*}
\end{linenomath*}
%
where, in this case, $\mu$ and $\mu'$ are as above, $\lambda=\lambda'=0$ and $\nu=0$, $\nu'=2$.
%
To retrieve the general solution to Eq. \ref{can_form} it is sufficient to substitute $(z^++\epsilon^+)/(H^+ +2 \epsilon^+)$ in place of $z^+$ in Eq. \ref{gen_sol_g}.
As the integration constants $\alpha$ and $\beta$ are specified through the imposition of boundary conditions, back-substituting $b^+ = (f_1+f_2)/2$ and $u^+ = -(f_1-f_2)/2i$ yields the solution in terms of $u^+$ and $b^+$.

% ####################################################################

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2014/09/25
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
%% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%%!PS-Adobe-3.0 EPSF-3.0
%%%BoundingBox: 19 19 221 221
%%%CreationDate: Mon Sep 29 1997
%%%Creator: programmed by hand (K)
%%%EndComments
%%
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%\documentclass[referee]{svjour3} 
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{natbib}
\linenumbers
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{epstopdf}
%\graphicspath{ {./images/results} }
\usepackage{color}
%\usepackage{ctable}
%\usepackage{epsfig}
%\usepackage{hyperref}
%\usepackage{soul}
%\usepackage{amssymb}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
 \journalname{Boundary-Layer Meteorology Research Note}

@@

\title{On the solution of katabatic flows with spatially varying eddy viscosity and diffusivity}

@@

%\authorrunning{Short form of author list} % if too long for running head
\author{M.~G.~Giometto \and R.~Grandi\and J.~Fang \and M.~B.~Parlange}

\institute{ M.~G.~Giometto \at School of Architecture, Civil and Environmental Engineering, \'Ecole Polytechnique F\'ed\'erale de Lausanne, Lausanne, Switzerland \\
		\email{marco.giometto@epfl.ch}
		\and
		R.~Grandi \at Mathematics Institute for Geometry and Applications, \'Ecole Polytechnique F\'ed\'erale de Lausanne, Lausanne, Switzerland \\
		 \and 
		 J.~Fang \at School of Architecture, Civil and Environmental Engineering, \'Ecole Polytechnique F\'ed\'erale de Lausanne, Lausanne, Switzerland \\
		 \and
		M.~B.~Parlange \at Dept. of Civil Engineering, Faculty of Applied Science, University of British Columbia, Vancouver, BC, Canada
}

@@

% ABSTRACT
\begin{abstract}
The Nieuwstadt closed form solution for the stationary Ekman layer is generalized for katabatic flows within the conceptual framework of the Prandtl model. 
The solution is valid for spatially-varying eddy viscosity and diffusivity (O'Brien type) and constant Prandtl number $(Pr)$. 
Momentum and buoyancy transfer coefficients are specified in accordance with Monin-Obukhov similarity.
The characteristics of the solution are discussed as a function of the dimensionless model parameters $Pr$ and $\hat{z}_0 \hat{N}^2 \hat{b}_s^{-1}$, where $\hat{z}_0$ is the hydrodynamic roughness length, $\hat{b}_s$ is the imposed surface buoyancy and $\hat{N}$ is the Brunt-V\"ais\"al\"a frequency.
For the considered range of such parameters, velocity and buoyancy profiles show significant variations in both phase and amplitude of extrema with respect to the classic constant-$K$ model and a more recent approximate solution based on the Wentzel-Kramers-Brillouin (WKB) theory.
Near-wall regions are characterized by relatively stronger surface momentum and buoyancy gradients, whose magnitude is inversely proportional to $Pr$. In addition, slope-parallel momentum and buoyancy fluxes are reduced, the low-level jet (LLJ) is further displaced toward the wall, and its peak velocity strongly depends on $\hat{z}_0 \hat{N}^2 \hat{b}_s^{-1}$.

%Simple approximate functions (based on the new solution) are provided, to relate the internal parameters of the system $(u_*, b_*)$ to the given external parameters.
%Universal functions are also provided to describe the variation of $ z_j$, $z_r$, and $\max{(u)}$, as a function of the dimensionless model parameter $z_0$.
\end{abstract}

\section{Introduction}
Slope winds are ubiquitous in nature and engineering, and they are of interest not only as a fundamental problem in itself, but also because of the important role they play over a broad range of scales and applications. 
On a local scale, for instance, they regulate local weather conditions, influencing atmospheric transport of momentum and scalars such as heat and humidity \citep{whiteman1990, Whiteman, Monti2002, Nylen2004b, Rotach2007, Lehner2015}. 
On a regional scale, they can be responsible for intense cyclonic vorticity in the middle and upper troposphere \citep{Parish1991, Parish1992, Parish1998}.
Persistent katabatic winds characterize the atmospheric boundary layer over Antarctica \citep{Chu1987, Renfrew2004, Renfrew2006a} and over glaciers \citep{Oerlemans1993, Oerlemans1994, pastex1994, Greuell1997, Smeets1997, Oerlemans1998, oerlemans1999, Smeets2000, Oerlemans2002}. Therefore an accurate characterization of such flows is an essential component of understanding and modeling of the weather and climate.
The problematic geometrical setup complicates measurements \citep{Oldroyd2015}, whereas the complex dynamics (e.g. turbulent intermittency, waves, Kelvin-Helmholtz instabilities, low-level jets) and the lack of a satisfactory similarity theory for such flows \citep{Nadeau2012} pose a challenge in terms of computational requirements for numerical modelers. In most cases the required resolution is in fact prohibitively costly \citep{Fedorovich2009, Fedorovich2009d, burkholder_2011}.
Because of this, conceptual models are still of great interest, and represent a valid tool for the characterization of such systems.

A cornerstone in the understanding of slope flows is represented by the classic Prandtl analytic model \citep{Prandtl}, afterwards extended to include Coriolis force \citep{Gutman1964}, external winds \citep{Lykosov_1972}, and surface variability \citep{Shapiro2007, Oldroyd2014,Zardi2015}, to name but a few.
The Prandtl model approximates the atmosphere in a Boussinesq sense and describes a steady stratified flow over a thermally perturbed unbounded planar sloping surface. It is summarized by the following system of ordinary differential equations:
\begin{linenomath*}
\begin{subequations}
	\begin{align}
	- \hat{N}^2 \hat{u}(\hat{z}) \sin{\alpha} & = [ \hat{K}_H \hat{b}_{\hat{z}} ]_{\hat{z}} \, , \\
	\hat{b}(\hat{z}) \sin{\alpha} & = [ \hat{K}_M \hat{u}_{\hat{z}} ]_{\hat{z}} \, ,
	\end{align}
	\label{Prandtl_eq}
\end{subequations}
\end{linenomath*}
where $\hat{(\cdot)}$ is used to denote a dimensional variable or parameter, $(\cdot)_{\hat{z}}$ denotes differentiation with respect to the slope-normal coordinate direction $\hat{z}$, $\hat{u}$ is the downslope velocity, $\hat{b} \equiv \hat{g} \hat{\theta}^{\prime} / \hat{\theta}_0$ is buoyancy, where $\hat{g}$ is the gravitational acceleration, $\hat{\theta}^{\prime}$ is the potential temperature perturbation (with respect to the background stably stratified environment characterized by the Brunt-V\"ais\"al\"a frequency $\hat{N}$), and $\hat{\theta}_0$ is a reference (constant) temperature; $\alpha$ is the slope angle and $\hat{K}_M$ and $\hat{K}_H$ denote the eddy viscosity and diffusivity (a model is used to parametrize turbulent fluxes of momentum and buoyancy, respectively).
Equations are defined in $\hat{z} \in \ [\hat{z}_0,\infty)$ with boundary conditions $\hat{u}(\hat{z}_0)=0$, $\hat{u}(\hat{z} \rightarrow \infty) = 0$, $\hat{b}(\hat{z}_0)=\hat{b}_s$ and $\hat{b}(\hat{z} \rightarrow \infty) = 0$ ($\hat{b}_s > 0$ for upslope flows, whereas $\hat{b}_s<0$ for downslope flows). 
Eqs. 1 state that the slope-parallel component of buoyancy is balanced by momentum flux divergence, and that slope-parallel buoyancy advection is balanced by buoyancy flux divergence.
The flow is assumed to be invariant in the along-slope direction and the model can be used to determine the slope-normal $(\hat{z})$ structure of velocity $\hat{u}(\hat{z})$ and buoyancy $\hat{b}(\hat{z})$.
The model is applicable far enough away from ridges and valleys that non-linear advection terms become small \citep{NAPPO1987}, and a balance between advection and diffusion (of both momentum and buoyancy) is achieved. This constraint is rather restrictive, as shown in \citet{Zardi2015}. 
The Prandtl constant-K solution reads 
\begin{linenomath*}
\begin{subequations}
	\label{Prandtl_sol}
	\begin{align}
		\hat{b} & = \hat{b}_s \exp{(-\hat{\sigma}_c \hat{z})} \cos{(\hat{\sigma}_c \hat{z})} \, , \\
		\hat{u} & = -\frac{\hat{b}_s}{N Pr} \exp{(-\hat{\sigma}_c \hat{z})} \sin{(\hat{\sigma}_c \hat{z})} \, ,
	\end{align}
\end{subequations}
\end{linenomath*}
where
\begin{eqnarray}
	\hat{\sigma}_c^2 \equiv \frac{\hat{\sigma_0}}{2\hat{K_H}} \, , \qquad \mathrm{and} \qquad
	\hat{\sigma}_0 \equiv \frac{\hat{N} \sin{(\alpha)}}{\sqrt{Pr}} \, .
\end{eqnarray}
The model is able to represent the low-level jet (LLJ) and the return flow region, key features of observed katabatic and anabatic flows. 
However, the constant-K solution is also known to be over-dissipative in the near-surface regions, and under dissipative above the LLJ regions \citep{Defant1949,Oerlemans1998, grisogono2001katabatic}. It is therefore not able to represent the observed strong surface gradients of temperature and momentum, and, in addition, the predicted wind speed typically decreases too rapidly away from the surface.
Simple variations in the eddy diffusivity profiles were introduced in a patched analytic solution by \citet{gutman_1983}, whereas more recently \citet{grisogono2001katabatic} considered general variations in the vertical structure of the eddy diffusivities, and derived  a patched global solution based on the WKB approximation \citep{Bender1979}.
The WKB solution to the Prandtl equations, valid to leading-order in the inner layer and to first-order in the outer layer, reads
\begin{subequations}
	\begin{align}
		\hat{f}_{in} & \sim  \exp{ \left( -(1 \pm i) (\hat{\sigma}_0/2)^{1/2} \int_0^{\hat{z}}{\hat{K}_H(z)^{-1/2}}{\mathrm{d}z}\right)}  &  \hat{z} \in [\hat{z}_0,\hat{h}] \, , \\
		\hat{f}_{out}&  \sim  [ \hat{K}_H(\hat{z}) / \hat{K}_H(\hat{h}) ]^{-1/4} \exp{ \left( -(1 \pm i) (\hat{\sigma}_0/2)^{1/2} \int_0^{\hat{z}}{\hat{K}_H(z)^{-1/2}}{\mathrm{d}z} \right)} & \hat{z} \in [\hat{h},\infty) \, ,
	\end{align}
	\label{WKB_sol}
\end{subequations}
where $\hat{f}_{in} \equiv \hat{b}_{in}+ i\hat{u}_{in}$ represents the inner-layer solution, and $\hat{f}_{out} \equiv \hat{b}_{out} + i\hat{u}_{out}$ is the outer-layer solution. $\hat{f}_{in}$ and $\hat{f}_{out}$ are patched at $\hat{z}=\hat{h}$, which separates the inner from the outer layer.
The WKB solution is able to account for additional dynamics while still retaining an elegant form.
However, WKB theory is only applicable when the model parameters $(\hat{K}_M,\hat{K}_H)$ vary more slowly than the solution $(\hat{u},\hat{b})$, and the validity of such a constraint for slope flows has been the subject of debate \citep{Grisogono2002}.

Here, a closed-form solution to the Prandtl-model equations is derived on a finite domain ($\hat{z} \in [\hat{z}_0,\hat{H}]$), valid for eddy viscosity and diffusivity coefficients that are modeled as a limited range of cubic polynomials, similar to what was proposed in \citet{O'Brien1970} for the planetary boundary layer. 
The derivation is a generalization of the solution proposed in \citet{Nieuwstadt1983a}, where the Ekman-layer equations were solved for the same form of momentum transfer coefficient.
Recall that the Ekman-layer equations can be reduced to the Prandtl equations after simple changes of variables \citep{Veronis1970}. 
The solution, expressed as a combination of Gaussian hypergeometric functions, represents an exact alternative to the WKB formulation for the chosen form of the eddy diffusivities. 
Its sensitivity to variations in the parameter space are investigated here within the Monin-Obukhov framework, to provide insight on the coupling between the velocity and buoyancy fields.

%sloping angle $\alpha \, \mathrm{(deg)}$ and hydrodynamic roughness length $z_0 \, \mathrm{(m)}$ parameters.

% ####################################################################
\section{Specification of $\hat{K}(\hat{z}$)}

%The characteristics of $\hat{K}(\hat{z})$ in katabatic flows were thoroughly justified in \citet{Grisogono2002}, whom outlined the following criteria that $\hat{K}(\hat{z})$ should honor:
%%
%\begin{enumerate}
%\item realizability condition: $\hat{K}(\hat{z}) \ge 0$;
%\item $\hat{K}(\hat{z}_0 < \hat{z} \ll \infty) = 0$;
%\item $\max(\hat{K}(\hat{z}))$, reached at $\hat{z}=\hat{h}$, must lie within the dynamic and thermal boundary layers;
%\item $\hat{h} > \max(2 \hat{z}_j, \hat{z}_{inv})$;
%\end{enumerate}
%%
%where $\hat{z}_j$ is the height of the LLJ, and $\hat{z}_{inv}$ is the height of the surface inversion layer (in general we have $\hat{z}_{inv} > \hat{z}_j$).

Here, eddy diffusivities are prescribed in line with the classic O'Brien's model \citep{O'Brien1970}, viz. 
%
\begin{equation}
	\hat{K}(\hat{z}) = \kappa \hat{u}_*\hat{z}(1-\hat{z}/\hat{H})^2  \qquad \hat{z} \in [\hat{z}_0,\hat{H}] \, ,
\end{equation}
%
where $\kappa$ is the von K\'arm\'an constant, $\hat{u}_* = \kappa (\hat{z} \hat{u}_{\hat{z}})|_{\hat{z}_0}$ is the friction-velocity, and $\hat{H}$ is the height of the domain, controlling both shape and magnitude of $\hat{K}$. 
The O'Brien model complies with the $\hat{K}$-requirements defined in \citet{Grisogono2002}, and has often been used in studies of stable boundary layers (see for instance \citet{pielke1984mesoscale} and \citet{Stull1988}). A generalized O'Brien model was also recently adopted in \citet{grisogono2001katabatic} to study katabatic flows.
In the original O'Brien's formulation $\hat{H}$ corresponds to the boundary layer depth; in this study $\hat{H}=3\hat{h}$ is evaluated iteratively under the constraint $\hat{h}=\hat{z}_r$, where $\hat{z}_r$ is the height of the peak velocity magnitude in the return flow region. 
Such a choice for $\hat{h}$ is based on a sensitivity analysis that was performed using direct numerical simulations to resolve katabatic flows over steep slopes ($\alpha \geq 15^{\circ}$).

Given the current lack of a rigorous similarity theory for katabatic flows, this study is restricted to the Monin-Obukhov framework, which is expected to yield acceptable approximations of transfer coefficients in the near-surface regions \citep{gutman_1983}.
Monin-Obukhov similarity is not expected to hold in the above-jet regions, where eddy viscosity and diffusivity coefficients are likely to depend on an additional set of parameters such as the sloping angle $(\alpha)$, the Brunt-V\"ais\"al\"a frequency $(\hat{N})$, and the imposed surface buoyancy $\hat{b}_s$ (or buoyancy flux). Nevertheless, the proposed solution is valid in a more general sense, and could easily be adapted in the future to a different $\hat{K}$-parameterisation. 
For instance, one could easily choose a different prefactor than $\kappa$, to make it depend on the model parameters.
Knowledge of $\hat{u}_*$ and $\hat{H}$ (or equivalently $\hat{h}$) allows one to specify $\hat{K}(\hat{z}) = \kappa \hat{u}_* \hat{z}(1-\hat{z}/\hat{H})^2$.

\section{The Analytic Solution}

For the combination
%
\begin{linenomath*}
\begin{equation}
   \hat{f}= \hat{b} - (i \hat{N} \sqrt{Pr}) \hat{u}
   \label{combinationVar}
\end{equation}
\end{linenomath*}
%
the system of Eqs. \ref{Prandtl_eq} is reduced to a complex ordinary differential equation (ODE) for the canonical variable $\hat{f}$:
%
\begin{linenomath*}
\label{ode_norm}
\begin{equation}
	\hat{f} = \left[ \frac{-i \hat{K}_M(\hat{z})}{\hat{N} \sin{(\alpha)} \sqrt{Pr}} \hat{f}_{\hat{z}} \right]_{\hat{z}} \qquad \hat{z} \in  \ [\hat{z}_0,\hat{H}] \, ,
	\label{ode}
\end{equation}
\end{linenomath*}
%
with boundary conditions $\hat{f}(\hat{z}_0)=\hat{b_s}$ and $\hat{f}(\hat{H}) = 0$.
Assigning a length, velocity and buoyancy scale $\hat{L}= \hat{u}_* \kappa (\hat{N} \sin{\alpha})^{-1}$, $\hat{U}=|\hat{b}_s| \hat{N}^{-1}$ and $\hat{B}=|\hat{b}_s|$ respectively, Eq. \ref{ode} reduces to:
%
\begin{linenomath*}
\begin{equation}
	f =  \frac{-i}{\sqrt{Pr}}\left[ K_M(z) f_z \right]_z \qquad  z \in [z_0,H] \, ,
	\label{ode_norm}
\end{equation}
\end{linenomath*}
%
where $K_M(z) = z(1-z/H)^2$ is the normalised eddy viscosity, $z=\hat{z}\hat{L}^{-1}$, $f = \hat{b} \hat{B}^{-1} + (i\sqrt{Pr})\hat{u}\hat{U}^{-1}$, $H=\hat{H}\hat{L}^{-1}$ and $z_0=\hat{z_0}\hat{L}^{-1}$, with boundary conditions $f(z_0)=-1$, $f(H) = 0$.
%A similar linear transformation was used in \citet{Nieuwstadt1983}, to solve for the Ekman equations.
%
The canonical form of Eq. \ref{ode_norm} reads:
%
\begin{linenomath*}
\begin{equation}
	\label{can_form}
	f_{zz}+P f_z+ Qf =0,
\end{equation}
\end{linenomath*}
%
where $P(z)=K_{M,z} / K_M$ and $Q(z)= (-i \, \sqrt{Pr}) /K_M$. Second, it is straightforward to show that rewriting Eq. \ref{can_form} for $y= \frac{z }{H}
$ results in
%if $f$ solves Eq. \ref{can_form} then $g(z^+):=f( (H^++2 \epsilon^+) z^+- \epsilon^+)$ solves:
%
\begin{linenomath*}
\begin{equation}
	\label{can_form_sec}
	f_{yy}+\widetilde{P} f_y+ \widetilde{Q} f =0,
\end{equation}
\end{linenomath*}
%
where $\widetilde{P}(y)=\gamma_y(y)/\gamma(y)$ and $\widetilde{Q}(y) = (-i \sqrt{Pr} H)/\gamma(y)$, with $\gamma(y)=y(y-1)^2$.
Eq. \ref{can_form_sec} is a second order ordinary differential equation with three regular singular points at $y = 0,1$ and $\infty$, as in \citet{Morse1954a}. 
This specific equation is known as the \emph{equation of Papperitz} and its general solution is:
%
\begin{linenomath*}
\begin{equation}
	f(y) = \alpha (1-y)^\mu \ _2F_1 (\mu, 1- \mu^{\prime}, 1+\mu-\mu^{\prime}, 1-y)
	+\beta (1-y)^{\mu^{\prime}} \ _2F_1 (\mu^{\prime}, 1- \mu, 1+\mu^{\prime}-\mu, 1-y) \, ,
	\label{gen_sol_g}
\end{equation}
\end{linenomath*}
%
where $\ _2F_1$ are \emph{Gaussian hypergeometric functions} 
%and 
%%
%\begin{equation}
%	\mu = -\frac{1}{2} + \frac{1}{2}\sqrt{1+4 i \sqrt{Pr} H} \,  , \qquad 
%	\mu^{\prime} = -\frac{1}{2} - \frac{1}{2}\sqrt{1+4 i \sqrt{Pr} H} \, ,
%\end{equation}
%
and $\mu$, $\mu'$ are the solutions to the quadratic equation

\begin{linenomath*}
\begin{equation*}
	x^2+x- i H \sqrt{Pr}=0.
\end{equation*}
\end{linenomath*}
